Stack Safety as a Security Property

### 2021-02-18 CheriApproaches.tex
Updated example to be much simpler

Either way the question becomes: should the variant trace be one with a policy active, or no? Let's
look at a concrete example.

\begin{verbatim}
main() {
  int[2] x = {0,0};
  int* y = &x+secret();
  f(&x);
}

f() {
  int** z = guess_y_addr();
  print(z**);
}
\end{verbatim}

Here {\tt main} has an array {\tt x} and saves a pointer to one of its elements, but which one is secret.
Then it passes the address to  {\tt f}, so that {\tt f} can access {\tt x} legally according to sharing.
{\tt x} has a tag that marks it permissible for anyone to access it, but {\tt f} has actually accessed {\tt x}
via reading the hidden information in {\tt y}. Suppose that our policy contains an error such that the access
of {\tt y} is successful - then {\tt f} always ends up printing 0 regardless, so we haven't violated
confidentiality. But we would if we printed {\tt y} outright or if the values in {\tt x[0]} and
{\tt x[1]} differed.

Consider the variant states that we can have. {\tt x} does not vary, but {\tt y} can point to {\tt x[0]},
{\tt x[1]}, or a location beyond {\tt x} entirely. In the first two cases, our output is 0,
so confidentiality is maintained. What do we do with that case? We can:

\begin{enumerate}
\item Recognize that the variant should fail-stop, and treat that as a violation
\item Recognize that the variant should fail-stop, and ignore this variant
\item Do not check for a variant fail-stop, and therefore likely get a violation
\end{enumerate}

I argue for option 2, because either some other variation that doesn't create an illegal pointer
will be a violation, or the particular trace has not leaked any information. In this case,
the error will be detected if we ever run the comparable case where {\tt x = {0,1}}, but if the compiler
somehow only produced traces where this issue is invisible, we need not consider it. Still, option 1
could be fine, since our policies are meant to prevent {\tt y} from being read in the first place.

Option 3 is clearly bad because if the contents of {\tt y} are used in some other way, its behavior
could be radically different from the primary trace. Whether it fails or not is virtually unrelated to
the initial potential violation. This seems very bad.

Considering the same situation in Cheri, we also have the possibility to get additional tag failures
by varying not just the address pointed to, but the permissions bits on a capability. But the fundamental
question remains the same: what to do with fail-stop variants.

>>>>

Lets consider a very simple machine with three addresses and an external feed of instructions
(instructions are not in memory.) An address can be marked either secret or public. Let's start with
the following configuration (top rows are addresses marked S or P, bottom are contents):

\begin{tabular}{| c | c | c |}
  \hline
  0(S) & 1(P) & 2(P) \\
  \hline
  0 & 0 & 0 \\
  \hline
\end{tabular}

We have a register \(r\), initialized to 0, and an output {\sc Out} that can be loaded to like a register.
We want to apply our confidentiality model by varying secret locations and determining if programs
differ in output between the original state and a variant. I'm using the simplified format from our
examples in the paper. Assume that loads use the address modulo three so that the step operation is
total. Consider the program:

\vspace{\abovedisplayskip}

\begin{tabular}{l l l}
  Line & Instr & Description \\
  \hline
  1 & \(r \leftarrow [0]\) & Load from address 0 \\
  2 & {\sc Out} \(\leftarrow [r]\) & Load from the resulting address and output it \\
  3 & \(r \leftarrow 0\) & Reset \(r\) to 0 \\
\end{tabular}

\vspace{\belowdisplayskip}

At the end of this program, 0 has been printed. In any variant when address 0 mods to a 1 or a 2, it
will also print 0. And since the final instruction clears \(r\), the final state is identical as well.
So we would not say that this program violates confidentiality of address 0, even though it was read.

Now suppose that we also make address 1 secret. The variation might vary both 0 and 1, so that the
program reads address 1 and outputs a non-zero result. Now, confidentiality is violated, because this
variation produces a different output. Suppose we consider some possible security properties to
prevent this from occurring. The property we come up with attaches to the load instruction and,
if the source register contains an address that is marked secret, fail-stops the machine. But we forget
to implement a similar rule for loading through immediate values, oops! So the load on line 1 never failstops.
Let's look at that second start state and the same program:

\begin{tabular}{| c | c | c |}
  \hline
  0(S) & 1(S) & 2(P) \\
  \hline
  0 & 0 & 0 \\
  \hline
\end{tabular}

Now the primary execution will load from address 0, then attempt to do so again and fail-stop. The variant
might instead try to read from a different location, but we are termination insensitive, so even if it reads
from 2 (successfully) the property holds. But what about this start state?

\begin{tabular}{| c | c | c |}
  \hline
  0(S) & 1(S) & 2(P) \\
  \hline
  2 & 0 & 0 \\
  \hline
\end{tabular}

Here the primary execution loads from 2 and prints 0. But a variant in which address 0 holds 0 or 1 will
load from those addresses, which would violate the policy and cause a failstop. So the question is, how do
we handle those variants? We can:

\begin{enumerate}
\item Recognize that the variant should fail-stop, and since the primary execution does not,
  treat that as a violation.
\item Recognize that the variant should fail-stop, and since we are termination insensitive,
  treat that as acceptable and look for another variant that actually produces different output.
\item Do not check whether the variant fail-stops. So when we vary address 0 to 1 and address 1 to 1,
  the program prints 1. Thus the property is violated.
\end{enumerate}

Options 1 and 2, in testing, will let us stop short and either move on to a different test or
declare a failure. Option 1 seems at odds with the notion of termination-insensitivity; so does option 3
in its own way, with the added downside that it's hard to predict whether a given case might fail. Option
2 is similar to only doing variants that maintain the legality of pointers. I've been leaning toward option
2, and I tend not to like option 3 because our variant run can show behavior that is infeasible in the
target machine.

This small example illustrates a scenario that can come up in HOPE and Cheri, when pointers and capabilities
being varied introduces fail-stop behavior in the variant trace when none was present in the primary. This
is an edge case, of course, because in any such scenario the program could instead print the pointer and thus
violate the property immediately. But we do need to decide how to approach it. We could imagine a policy that
prevents printing the pointer directly (via taint-tracking, say) but does not prevent dereferencing it.
